---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

Package needed:
```{r}
library(ggplot2)
library(knitr)
library(dplyr)
library(lubridate)
```
```{r setoptions, echo = TRUE}
opts_chunk$set(echo = TRUE)
```

File zip verification. If the zip file exist, unzip it and read the data:
```{r}
if(file.exists("activity.zip")){
	activity <- read.csv(unzip("activity.zip"))
}
```

Transforme Dates in Date formate
```{r}
activity$date <- ymd(activity$date)
```

## What is mean total number of steps taken per day?
Calculate the total number of steps taken per day
(ignore the missing values)
```{r}
steps <- activity %>%
         group_by(date) %>%
         summarize(steps = sum(steps))
```

Plot the histogramm of the total number of steps taken per day

```{r}
qplot(steps, data = steps, xlab = "Total steps per day", bins = 10)
```

The following code:
```{r}
mean <- mean(steps$steps, na.rm = TRUE)
median <- median(steps$steps, na.rm = TRUE)
total <- sum(steps$steps, na.rm = TRUE) 
```
allows to obtain the mean, `r mean`, and median, `r median`, values of the total number of steps taken per day and the total number of steps, `r total`.

## What is the average daily activity pattern?

Calculate the total number of steps in each 5-minutes interval across all days:
```{r}
stepsInterval <- activity %>%
         filter(!is.na(steps)) %>%
         group_by(interval) %>%
         summarize(steps = sum(steps))
```

A time series plot of the 5-minute interval and the average number of steps: 
```{r}
plot(stepsInterval$interval, stepsInterval$steps, type = "l", xlab = "5-minute Interval", ylab = "Average number of steps", main = "Daily activity pattern")
```

The following code:
```{r}
max <- stepsInterval[which.max(stepsInterval$steps),]

```
allows to obtain the interval contains the maximum number of steps on average across all the days in the dataset, which is: `r max[1]`

## Imputing missing values

The  total number of missing values (total number of rows with NAs):
```{r}
totalNA <- sum(is.na(activity$steps))
```
is `r totalNA`.

The mean value of steps for a 5-minute interval is used to fill in the NAs in original dataset.

```{r}
stepsMean <- activity %>%
         filter(!is.na(steps)) %>%
         group_by(interval) %>%
         summarize(steps = mean(steps))

notNA <- activity

for (i in 0:60){
    for (j in 1:288){
      k<- i*288+j
      if (is.na(notNA$steps[k])){
          notNA$steps[k] <- stepsMean$steps[j]  
      }
    }
}
```
Calculate the total number of steps taken per day

```{r}
stepsNA <- notNA %>%
           group_by(date) %>%
           summarize(steps = sum(steps))
```
Plot the histogramm of the total number of steps taken per day

```{r}
qplot(steps, data = stepsNA, xlab = "Total steps per day", bins = 10)
```

The following code:
```{r}
meanNA <- mean(stepsNA$steps, na.rm = TRUE)
medianNA <- median(stepsNA$steps, na.rm = TRUE)
totalNA <- sum(stepsNA$steps, na.rm = TRUE)
stepsDiffer <- totalNA - total
```
allows to obtain the mean, `r meanNA`, and median, `r medianNA`, values of the total number of steps taken per day with imputed missing values. The mean do not differ from the mean value without impute missing values, `r mean`, as we use mean for replace NAs. But addition values increased the  median in comparing with calculated without imputed missind values, `r median`.
Total number of steps `r totalNA` increases in comparison with number of steps without impute, `r total` on `r stepsDiffer` steps.

## Are there differences in activity patterns between weekdays and weekends?

Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Note! I use the german version, for english version "Samstag" and "Sonntag" in code should be replaced with "Saturday" and "Sunday"

```{r}
notNA$Day <- ifelse(weekdays(notNA$date) == "Samstag" | weekdays(notNA$date) == "Sonntag", "weekend", "weekday")
notNA$Day <- as.factor(notNA$Day)
```

Calculate the average steps in the 5-minute interval for weekday and weekend

```{r}
intervalDay <- notNA %>%
    group_by(interval, Day) %>%
    summarise(steps = mean(steps))
```

Use ggplot for making the time series of the 5-minute interval for weekday and weekend, and compare the average steps:

```{r}
ggplot(intervalDay, aes(interval, steps, color = Day)) + geom_line() +facet_wrap(~Day, ncol = 1, nrow=2)
```

Plots above shows less activity on early mornings on weekends but more activity during the day long.